---
title       : Developing Data Products Course Project
subtitle    : Predicting Iris Species
author      : Luke Chen
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Predicting Iris Species

* Created app that predicts Iris species using petal length and width
* Using decision tree model to fit data
* Give most likely species and probability

--- .class #id 

## Model fit

Model is first fitted with dataset iris

```{r,echo=FALSE}
  library(rpart)
  library(rpart.plot)
  fit <- rpart(Species ~ Petal.Length+Petal.Width, data = iris)
  fit
```

--- .class #id 


## Example

For example, if petal length is 3 cm and petal width is 1cm, it outputs the splitting criteria and prediction

```{r simple-plot, fig.height=3,echo=FALSE}
  library(rpart)
  library(rpart.plot)
  fit <- rpart(Species ~ Petal.Length+Petal.Width, data = iris)
  rpart.plot(fit)
  df<-data.frame(Petal.Length=3,Petal.Width=1)
  predict(fit,df,"class")
  predict(fit,df)
```

--- .class #id

## The End

Thanks for your time. Shiny code can be found in 


https://github.com/lukeyanchen/DevelopingDataProducts
