---
title: "Statistical Inference Course Project Part 1"
author: "Luke Chen"
output: pdf_document
toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.pos = 'H')
```

## Overview

In this project, we will investigate the exponential distribution and compare it with the Central Limit Theorem. We will investigate the distribution using a thousand simulations. Primarily, we are going to:

1. Show the sample mean and compare it to the theoretical mean of the distribution.
2. Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.
3. Show that the distribution is approximately normal.

## Simulation

The exponential distribution can be simulated in R with **rexp(n, lambda)** where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. We are going to set lambda = 0.2 for all of the simulations. We will investigate the distribution of averages of 40 exponentials and run a thousand simulations.

```{r}
#setting seed for replication
set.seed(123)

#setting lambda to 0.2
lambda<- 0.2

#setting n to 40
n<- 40

# number of simulations
simulation<- 1000

#running simulations
sim_mean<- NULL
for (i in 1 : 1000) sim_mean = c(sim_mean, mean(rexp(n,lambda)))
```

## 1. Sample Mean vs. Theoretical Mean

```{r}
# sample mean
mean(sim_mean)

#theoretical mean
1/lambda
```

The sample mean is very close to the theoretical mean. The comparison is graphically shown in *Figure 1* under the appendix.


## 2. Sample Variance vs. Theoretical Variance

```{r}
# sample variance
var(sim_mean)

# theoretical variance
(1/lambda)^2/n
```

Theoretical SD is $\frac{1/\lambda}{\sqrt{n}}$ and so theoretical variance is $\frac{1/\lambda^2}{n}$


The sample variance is smaller than theoretical variance but is very close.
Graph below will show a comparison of sample SD and theoretical SD as variance is just the squared of SD. The comparison is graphically shown in *Figure 2* under the appendix.


## 3. Is distribution normal?

We show that the distribution of a large collection of averages of 40 exponentials (*Figure 4*) is much closer to the normal distribution than the distribution of a large collection of random exponentials (*Figure 3*).
We have also imposed a normal distribution of mean $5$, sd $\sqrt{0.625}$ as a reference. Based on the CLT, the distribution of averages of exponentials will get close and closer to the normal distribution as no. of simulation increases.


## Appendix

```{r,fig.align='center',fig.cap=" ",fig.height=3.5}
library(ggplot2)
cuts1 <- data.frame(Mean="Sample", vals=c(mean(sim_mean)))
cuts2 <- data.frame(Mean="Theoretical", vals=c(1/lambda))
cuts<-rbind(cuts1,cuts2)
qplot(sim_mean, xlab="Simulated means", 
main = "Sample Vs Theoretical Mean",binwidth=.1, fill=I("grey"),col=I("black"))+
geom_vline(data=cuts, aes(xintercept=vals, linetype=Mean,colour = Mean),show.legend = TRUE)

cuts1 <- data.frame(SD="Sample", 
                    vals=c(mean(sim_mean)-sqrt(var(sim_mean)),
                    mean(sim_mean)+sqrt(var(sim_mean))))
cuts2 <- data.frame(SD="Theoretical",
                    vals=c(1/lambda-1/lambda/sqrt(n),
                           1/lambda+1/lambda/sqrt(n)))
cuts<-rbind(cuts1,cuts2)
qplot(sim_mean, xlab="Simulated means", main = "Sample Vs Theoretical SD",binwidth=.1, 
fill=I("grey"),col=I("black"))+
geom_vline(data=cuts, aes(xintercept=vals, linetype=SD,colour = SD),show.legend = TRUE)

hist(rexp(1000,1/lambda),xlab="",main="Distribution of large collection of random exponentials")
qplot(sim_mean, xlab="Simulated means", 
main = "Sample Distribution vs Normal Distribution",
binwidth=.1, fill=I("grey"), col=I("black"))+
stat_function( fun = function(x, mean, sd, n, bw){ 
      dnorm(x = x, mean = mean, sd = sd) * n * bw
    }, args = c(mean = 5, sd = sqrt(0.625), n = 1000, bw = 0.1), col="red")
```

